---
title: "A study of influences on charitable giving for conservation of forests (Aug 2017)"
author: "Anusha Sridhara"
date: "11/24/2021"
output: html_document
---

## Introduction

Batavia and Chelsea (2017) from Andrews Forest LTER Site, conducted an online survey, where 1331 participants across United States responded to questionnaires about social desirability, moral inclusivity and donation set-up. The survey asked respondents to indicate their political willingness and whether they consider themselves as an environmentalist or otherwise. This report analyzes likelihood of association between people's political affiliation and their willingness to donate for conservation of forests with the help of a non-parametric test, Pearson's Chi-Squared test.

## Data Analysis
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE)
#Attaching the packages
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(dplyr)
library(kableExtra)
```

```{r, include=FALSE}
#Reading in the data
survey <- read_csv(here("data", "conservation_survey.csv")) 
```

```{r, include=FALSE}
#cleaning up column headings
survey_clean <- clean_names(survey)
```

```{r, include=FALSE}
#Creating a subset with only data from political affiliation and environmentalist
paff_envt <- survey_clean %>% # create new file paff_envt from data survey_clean and
  select(env_id, pol_or) # include only columns env_id and pol_or
```

```{r, include=FALSE}
#Simplifying data frame with specified political affiliation and if respondents are environmentalists or not
paff_envt1 <- paff_envt %>%  # Create new file paff_envt1 from paff_envt and add new columns simplifying data using mutate 
  mutate(political_affiliation = case_when(
    pol_or %in% c(1,2,3) ~ "Conservative",
    pol_or %in% c(5,6,7) ~ "Liberal")) %>% 
  drop_na() %>% 
  mutate(environmentalist = case_when(
    env_id %in% c(1) ~ "Yes",
    env_id %in% c(2) ~ "No",
    env_id %in% c(3) ~ "Unsure"
  )) # renamed values of political affiliation and environmentalism 
```


```{r, include=FALSE}
#Calculating total counts by political affiliation and environmentalist
paff_envt2 <- paff_envt1 %>% # Create file paff_envt2 from paff_envt1
  count(political_affiliation, environmentalist) # and count the number of responses

#Presenting the counts in tidy format using janitor::tabyl
paff_envt_counts <- paff_envt1 %>% 
  janitor::tabyl(political_affiliation, environmentalist)

paff_envt_counts
```


```{r, include=TRUE}
#Calculating proportions by row
pol_aff_proportions <- paff_envt_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits=2) %>% 
  janitor::adorn_ns() %>% 
  kable(caption = "*Table 1: Total survey counts and proportions of responses identifying self as environmentalist by political affiliation*") %>% 
  kable_styling(bootstrap_options = "striped", full_width = TRUE)

pol_aff_proportions
```

```{r, include=FALSE}
#Converting to a contingency table - column to row names
pol_aff_ct <- paff_envt_counts %>% 
  column_to_rownames(var = "political_affiliation") %>% 
  kable(caption = "Table 1: Total counts of survey responses identifying self as environmentalist by political affiliation") %>% 
  kable_styling(bootstrap_options = "striped", full_width = TRUE)
pol_aff_ct
```

```{r, include=FALSE}
#Performing Chi-Square test to assess the association between political affiliation and considering self as an environmentalist
pol_aff_cst <- chisq.test(paff_envt_counts)
pol_aff_cst
```

```{r, include=FALSE}
# Creating a tidy version of the contingency table 
pol_aff_cst_tidy <- broom::tidy(pol_aff_cst)
pol_aff_cst_tidy
```

## Pearson's Chi-Squared Test Interpretation

A Pearson's Chi-squared test was performed on data of 1331 respondents. The results have revealed that there is a significant association between one's political affiliation (whether one identifies self as a liberal or a conservative) and likelihood of identifying themselves as an environmentalist or otherwise ($\chi$^2^(`r pol_aff_cst_tidy$parameter`) = `r round( pol_aff_cst_tidy$statistic, 2)`, *p* = `r pol_aff_cst_tidy$p.value`. The results could therefore also help in predicting whether a person's political affiliation and their tendency to identify as an environmentalist or not, would in turn have an effect on their likelihood of making charitable contributions towards conservation of forests. In other words, political affiliation of a person does affect their likelihood to make charity towards conservation of forests.


## Data Citation
Batavia, C. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c (Accessed 2021-11-28).










